Blame and Coercion: Together Again for the Second Time

### 2019-10-16 paper.tex

%% Abstract
%% Note: \begin{abstract}...\end{abstract} environment must come
%% before \maketitle command
\begin{abstract}
  \todo[inline]{Rewrite this abstract based on the comparision table. -Jeremy}
  Gradual Typing is great. Implementing gradually typed with blame
  tracking and space efficiency is tricky. There exist two technique
  to do this: coercion and threesome. Coercion is easy to understand,
  and easy enough to implement, but difficult to reason about
  formally. Threesome is hard to understand, easy to implement, and
  easy to reason about formally. We propose hypercoercion, which is
  easy to understand, as easy to implement as coercion, and easy to
  reason about formally.
  \todo[inline]{Spell it ``hypercoercion'' without the hyphen. -Jeremy}
\end{abstract}

@@

@@

\citet{siek2009exploring} decorate \citet{herman2010space}'s coercions with 
labels to restore 
blame-tracking. They also propose that there are four blame 
strategies for their coercion:
$ \{Lazy, Eager\} \times \{D, UD\} $. 
Eager strategies detect more potential type errors, 
but blame more programs.
$ D $ and $ UD $ assign blame labels differently.
They prove partially that their \lazyUD{} coercion is correct by showing 
that it simulates the Blame Calculus in \cite{wadler2009well}. It is unknown, 
however, whether the simulation also works in the other direction.

\citet{siek2010threesomes} proposes another approach to combine blame 
tracking and space efficiency. Their solution represents casts with threesomes, 
a novel cast representation.
They show that their calculus bisimulates the \lazyUD{} coercion 
calculus in \citet{siek2009exploring} and the Blame Calculus in 
\citet{wadler2009well}. They also prove that threesome is space-efficient as 
well.

@@

Following \citet{siek2010threesomes}, \citet{garcia2013calculating} develops
threesome for other blame strategies. After studying in depth the relation 
between coercion and threesome,
he claims that normal coercion with labels is easy to understand but hard to 
implement and hard to reason about, and that 
threesome with labels, however, is easy to implement and easy to reason about 
but hard to understand. 
His claim is mostly affirmed by the group of people who develop threesome 
(\citet{siek2015blame}).
His threesome calculi are equivalent to the coercion calculi in 
\citet{siek2009exploring} because the former ones are derived from the latter 
ones.

@@

@@

\lazyD{} HC constructors are in \figref{fig:HC-D}.

$ c_1 \fatsemi^\ell c_2 = c' $ composes two hypercoercions. The target of $ 
c_1 $ and the source of $ c_2 $ might be different, in which case $ \ell $ must 
be a label.
$ m_1 \fatsemi^\ell (m_2,t) = (m',t') $ composes a middle with a 
pair of a middle and a tail. Similarly, the target of $ m_1 $ and the source of 
$ m_2 $ might be different, in which case $ \ell $ must be a label.
$ seq(c_1,c_2) $ composes two coercions where the target type of $ c_1 $ is 
equal to the source type of $ T_2 $.
$ id(T) $ constructs an identity coercion of $ T $ with the help of $ id(P) $.
$ cast(T_1,l,T_2) $ constructs a coercion from a source type, a label, and a 
target type.

\begin{proposition}[\lazyD\ hypercoercions is a Monoid]
	For all $c : T_1 \Longrightarrow T_2$,
	$c_1 : T_1 \Longrightarrow T_2$,
	$c_2 : T_2 \Longrightarrow T_3$, and
	$c_3 : T_3 \Longrightarrow T_4$,
	\begin{enumerate}
		\item $seq(id(T_1),c) = c$,
		\item $seq(c,id(T_2)) = c$, and
		\item $seq(seq(c_1, c_2), c_3) = seq(c_1, seq(c_2, c_3))$.
	\end{enumerate}
\end{proposition}

@@

\subsection{Relating All Surely \lazyD{} Cast Representations}

\begin{proposition}[Strong Bi-simulation among $ \mathcal{S}(\cdot) $]
	\label{thm:CEKS-bisim}
	If 
	$ C_1 $ and $ C_2 $ are surely \lazyD,
	$ s_1, s_2 \in S(C_1) $ and $ s_3 \in S(C_2) $,
	$ s_1 \approx s_3 $,
	$ \judgeSreduce{C_1}{s_1}{s_2} $,
	then there exists an $ s_4 $ such that
	\begin{itemize}
	\item $ s_2 \approx s_4 $ and
	\item $ s_3 \longrightarrow_{S(C_2)} s_4 $
	\end{itemize}
\end{proposition}
\begin{proof}
	The key ideas of this proof are undoing sequencing with the property (2) of 
	surely \lazyD{} CR, and handling all possibly uses of $ 
	cast(T,l,T) $ with property (3)-(11).
\end{proof}

\begin{proposition}[Equivalence of Surely \lazyD{} Cast Representations]
	\label{thm:surely-lazyD-eqv}
	If $ \judgetype{\emptyset}{e}{T} $, $ o : T $, and $ C_1 $ and $ C_2 $ 
	are surely \lazyD cast representations,
	\[
	e \Downarrow_{S(C_1)} o \; \text{if and only if} \; 
	e \Downarrow_{S(C_2)} o
	\]
\end{proposition}
\begin{proof}
	Immediately from Proposition~\ref{thm:CEKS-bisim}.
\end{proof}

\subsection{Relating All Surely \lazyD{} Cast Representations to $\mathcal{D}$}

So far, we have related all surely \lazyD{} CRs, i.e. they are equivalent to 
each other. But this does not mean they respect the semantics of the \lazyD{} 
Blame Calculus. We solve prove this part in this subsection.

The state bisimilarity here ($\eqvDS{C}{s}{s}$) is mostly derived from the 
definitions of $ \mathcal{D} $. Exceptions include the bisimilarity relations 
between continuations, between values, and from continuations to 
pre-continuations.

\todo[inline]{bisimilarity}

@@

\subsection{\lazyD\ hypercoercion Respect $ \mathcal{D} $}
\label{sec:LDHC-correct}

To show that \lazyD\ HC is correct, we first show it is surely \lazyD, then 
uses the Theorem~\ref{thm:surely-lazyD-correct} to prove its correctness.

@@

\begin{theorem}[\lazyD{} hypercoercion Respect $ \mathcal{D} $]
	If $ \judgetype{\emptyset}{e}{T} $ and $ o : T $ 
	\[
	\judgeDBeval{e}{o} \; \text{if and only if} \; 
	\judgeSeval{H}{e}{o}
	\]
\end{theorem}
\begin{proof}
	Immediately from Lemma~\ref{thm:hc-surely-lazyD} and 
	Theorem~\ref{thm:surely-lazyD-correct}.
\end{proof}

>>>>

%% Abstract
%% Note: \begin{abstract}...\end{abstract} environment must come
%% before \maketitle command
\begin{abstract}
\todoKC{rewrite this abstract based on the comparision table. -Jeremy}
  Gradual Typing is great. Implementing gradually typed with blame
  tracking and space efficiency is tricky. There exist two technique
  to do this: coercion and threesome. Coercion is easy to understand,
  and easy enough to implement, but difficult to reason about
  formally. Threesome is hard to understand, easy to implement, and
  easy to reason about formally. We propose hypercoercion, which is
  easy to understand, as easy to implement as coercion, and easy to
  reason about formally.
\todo[inline]{Spell it ``hypercoercion'' without the hyphen. -Jeremy}
\end{abstract}

@@

\todo[inline]{I see lots of mistakes regarding when to make words
  plural (whether they should have an ``s'' at the end). Please review
  the rules of English regarding plurals. -Jeremy}

@@

\citet{siek2009exploring} decorate \citet{herman2010space}'s coercions with 
labels to restore 
blame-tracking. They also propose that there are four blame 
strategies for their coercion:
$ \{Lazy, Eager\} \times \{D, UD\} $. 
Eager strategies detect more potential type errors, 
but blame more programs.
$ D $ and $ UD $ assign blame labels differently.
They prove their \lazyUD{} coercion simulates the Blame Calculus in 
\cite{wadler2009well}. It is unknown, however, whether the simulation in the 
other direction also works.

\citet{siek2010threesomes} proposes another approach to combine blame 
tracking and space efficiency. Their solution represents casts with threesomes, 
a novel cast representation. Threesome is proved space-efficient as well.
They prove that their calculus bisimulates the \lazyUD{} coercion 
calculus in \citet{siek2009exploring} and the Blame Calculus in 
\citet{wadler2009well}.

@@

\citet{garcia2013calculating} develops
threesome for other blame strategies. After studying in depth the connection 
between coercion and threesome,
he claims that normal coercion with labels is easy to understand but hard to 
implement and hard to reason about, and that 
threesome with labels, however, is easy to implement and easy to reason about 
but hard to understand. 
His claim is mostly affirmed by the group of people who develop threesome 
(\citet{siek2015blame}).
His threesome calculi are equivalent to the coercion calculi in 
\citet{siek2009exploring} because the former ones are derived from the latter 
ones.

@@

\todoKC{change the reduction rules of case}

@@

\figref{fig:HC-D} defines the constructors of \lazyD{} HC.

\todoKC{check consistency before using middle composition.}
$ c_1 \fatsemi^\ell c_2 = c' $ composes two hypercoercions. The target of $ 
c_1 $ and the source of $ c_2 $ might be different, in which case $ \ell $ must 
be a label.
$ m_1 \fatsemi^\ell (m_2,t) = (m',t') $ composes a middle with a 
pair of a middle and a tail. Similarly, the target of $ m_1 $ and the source of 
$ m_2 $ might be different, in which case $ \ell $ must be a label.
$ seq(c_1,c_2) $ composes two coercions where the target type of $ c_1 $ is 
equal to the source type of $ T_2 $.
$ id(T) $ constructs an identity coercion of $ T $ with the help of $ id(P) $.
$ cast(T_1,l,T_2) $ constructs a coercion from a source type, a label, and a 
target type.

\begin{proposition}[\lazyD\ hypercoercion is a monoid]
	For all $c : T_1 \Longrightarrow T_2$,
	$c_1 : T_1 \Longrightarrow T_2$,
	$c_2 : T_2 \Longrightarrow T_3$, and
	$c_3 : T_3 \Longrightarrow T_4$,
	\begin{enumerate}
		\item $seq(id(T_1),c) = c$,
		\item $seq(c,id(T_2)) = c$, and
		\item $seq(seq(c_1, c_2), c_3) = seq(c_1, seq(c_2, c_3))$.
	\end{enumerate}
\end{proposition}

@@

\subsection{All Surely \lazyD{} and Monoidic Cast Representations Respect 
$\mathcal{D}$}

We use bisimulation to proof equivalence between the two machines.
The state bisimilarity ($\eqvDS{C}{s}{s}$) is mostly derived from the 
definitions of $ \mathcal{D} $. Exceptions include the bisimilarity relations 
between continuations, between values, and from continuations to 
pre-continuations. We omit subscripts when there is no ambiguity.

\todoKC{write down the exceptional bisimilarity relations}

@@

\subsection{\lazyD\ Hyper-coercion Respect $ \mathcal{D} $}
\label{sec:LDHC-correct}

@@

\begin{theorem}[\lazyD{} hypercoercion Respect $ \mathcal{D} $]
	If $ \judgetype{\emptyset}{e}{T} $ and $ o : T $ 
	\[
	\judgeDBeval{e}{o} \; \text{if and only if} \; 
	\judgeSeval{H}{e}{o}
	\]
\end{theorem}
\begin{proof}
	Immediately from Lemma~\ref{thm:surely-monoidic-eval},
	Lemma~\ref{thm:hc-surely-lazyD}, and 
	Proposition~\ref{thm:hc-monoid}.
\end{proof}

\subsection{Extra Theorems}

In these section we show some extra theorems that might be useful in proving 
the correctness of other cast representations.

\begin{lemma}[Strong Bi-simulation among $ \mathcal{S}(\cdot) $]
	\label{thm:CEKS-bisim}
	If 
	$ C_1 $ and $ C_2 $ are surely \lazyD,
	$ s_1, s_2 \in S(C_1) $ and $ s_3 \in S(C_2) $,
	$ s_1 \approx s_3 $,
	$ \judgeSreduce{C_1}{s_1}{s_2} $,
	then there exists an $ s_4 $ such that
	\begin{itemize}
	\item $ s_2 \approx s_4 $ and
	\item $ s_3 \longrightarrow_{S(C_2)} s_4 $
	\end{itemize}
\end{lemma}
\begin{proof}
	The key ideas of this proof are undoing sequencing with the property (2) of 
	surely \lazyD{} CR, and handling all possibly uses of $ 
	cast(T,l,T) $ with property (3)-(11).
\end{proof}

\begin{proposition}[Equivalence of Surely \lazyD{} Cast Representations]
	\label{thm:surely-lazyD-eqv}
	If $ \judgetype{\emptyset}{e}{T} $, $ o : T $, and $ C_1 $ and $ C_2 $ 
	are surely \lazyD\ cast representations,
	\[
	e \Downarrow_{S(C_1)} o \; \text{if and only if} \; 
	e \Downarrow_{S(C_2)} o
	\]
\end{proposition}
\begin{proof}
	Immediately from Lemma~\ref{thm:CEKS-bisim}.
\end{proof}
